### 2016-06-29 GPCE2016/sections/CaseStudy.tex

\noindent \texttt{ExpPC} is the new expression interface supporting
\texttt{print} and \texttt{collectLit} operations; \texttt{LitPC} and
\texttt{AddPC} are the extended variants. Notice that except for the routine of
\textbf{extends} clauses, no glue code is required. Return types of
\texttt{e1,e2} must be refined to \texttt{ExpPC}.

Note that the code for instantiation is generated by \mixin.
Creating a simple expression of type \texttt{ExpPC} is
as simple as:
\begin{lstlisting}
ExpPC e8 = AddPC.of(LitPC.of(3), LitPC.of(4));
\end{lstlisting}
\noindent Without our approach, tedious instantiation code would need
to be defined manually.

@@

Smalltalk and Dart offer method cascading and avoid the need for fluent setters.
It is achieved at the price of introducing additional syntax and 
intrinsically relies on an imperative setting. 
Our approach supports both fluent setters and (functional) 
fluent withers.

@@

\noindent The \mixin annotation allowed us to reduce the interfaces/classes used
in Bono et al.'s implementation by 21.4\% (from 14 to 11). The
reductions were due to the replacement of instantiation classes with
generated \texttt{of} methods. The number of source lines of code 
(SLOC\footnote{Only Java code is counted, excluding comments and blank lines. })
was reduced by 40\% due to both the removal of instantiation overhead and
generation of getters/setters. %\bruno{explain why? we generate code for getters/setters?}.
To ensure a fair comparison, we used the same coding style as Bono et al.'s.

@@

Our refactoring applies the pattern presented in
Section~\ref{subsec:ep} to the existing Mumbler code base to improve
its modularity and extensibility. Using the refactored code base it
becomes possible to add new operations modularly, and to support
independent extensibility. We add one more
operation \texttt{print} to both the original and the refactored code
base. In the original code base the pretty printer is added
non-modularly by modifying the existing code. In the refactored code
base the pretty printer is added modularly. 
Although the code in the refactored version is slightly increased (by 2.4\% SLOC), the
modularity is greatly increased, allowing for improved reusability and maintainability.
We list the results in the following table:

>>>>

\noindent \texttt{ExpPC} is the new expression interface supporting
\texttt{print} and \texttt{collectLit} operations; \texttt{LitPC} and
\texttt{AddPC} are the extended variants. Notice that except for the routine of
\textbf{extends} clauses, no glue code is required. Return types of
\texttt{e1,e2} must be refined to \texttt{ExpPC}.
Creating a simple expression of type \texttt{ExpPC} is
as simple as:
\begin{lstlisting}
ExpPC e8 = AddPC.of(LitPC.of(3), LitPC.of(4));
\end{lstlisting}
\noindent Without Classless Java, tedious instantiation code would need
to be defined manually.

@@

Languages like Smalltalk and Dart offer method cascading and avoid the need for fluent setters.
This is achieved at the price of introducing additional syntax and 
intrinsically relies on an imperative setting. 
Our approach supports both fluent setters and (functional) 
fluent withers.

@@

\noindent The \mixin annotation reduced the interfaces/classes used
in Bono et al.'s implementation by 21.4\% (from 14 to 11), due to the replacement of instantiation classes with
generated \texttt{of} methods. The number of source lines of code 
(SLOC)
was reduced by 40\% due to both the removal of instantiation overhead and
generation of getters/setters. %\bruno{explain why? we generate code for getters/setters?}.
%To ensure a fair comparison, we used the same coding style as Bono et al.'s.

@@

Our refactoring applies the pattern presented in
Section~\ref{subsec:ep} to the existing Mumbler code base to improve
its modularity and extensibility. Using the refactored code base it
becomes possible to add new operations modularly, and to support
independent extensibility. We add one more
operation \texttt{print} to both the original and the refactored code
base. In the original code base the pretty printer is added
non-modularly by modifying the existing code. In the refactored code
base the pretty printer is added modularly. 
Although the code in the refactored version is slightly increased (by 2.4\% SLOC), the
modularity is greatly increased, allowing for improved reusability and maintainability:

### 2016-06-29 GPCE2016/sections/RelatedWork.tex

\paragraph{Multiple inheritance in \objectoriented languages}
Many authors have argued in favor or against multiple inheritance.
Multiple \emph{implementation}
inheritance is very expressive, but difficult to model and implement, and can
cause difficulties (including the famous diamond (fork-join)
problem~\cite{bracha90mixin,Sak89dis}, conflicting methods, etc.) in reasoning about
programs. To conciliate the need for expressive power and simplicity, many
models have been proposed, including C++ virtual inheritance,
mixins~\cite{bracha90mixin}, traits~\cite{scharli03traits}, and hybrid models
such as CZ~\cite{malayeri2009cz}.  They provide novel programming architecture
models in the OO paradigm. In terms of restrictions set on these models, C++
virtual inheritance aims at a relative general model; the mixin model adds some
restrictions; and the trait model is the most restricted one (excluding
state, instantiation, etc).

%\paragraph{C++ Approach.}
C++ has a general solution to multiple inheritance by
virtual inheritance, dealing with the diamond problem by keeping only
one copy of the base class~\cite{ellis1990annotated}. However C++
suffers from the object initialization problem~\cite{malayeri2009cz}.
It bypasses constructor calls to virtual superclasses, which can
cause serious semantic errors. In our approach, the \mixin
annotation has full control over object initialization,
and the mechanism is transparent to users. Moreover, customized factory
methods are also allowed: if users are not satisfied with the default
generated \texttt{of} method, they can implement their own.

>>>>

\paragraph{Multiple inheritance in \objectoriented languages}
Many authors have argued in favor or against multiple inheritance.
Multiple \emph{implementation}
inheritance is very expressive, but difficult to model and implement, and can
cause difficulties (including the famous diamond %(fork-join)
problem~\cite{bracha90mixin,Sak89dis}, conflicting methods, etc.) in reasoning about
programs. To conciliate the need for expressive power and simplicity, many
models have been proposed, including C++ virtual inheritance,
mixins~\cite{bracha90mixin}, traits~\cite{scharli03traits}, and hybrid models
such as CZ~\cite{malayeri2009cz}.  They provide novel programming architecture
models in the OO paradigm. In terms of restrictions set on these models, C++
virtual inheritance aims at a relative general model; the mixin model adds some
restrictions; and the trait model is the most restricted one (excluding
state, instantiation, etc).

%\paragraph{C++ Approach.}
C++ has a general solution to multiple inheritance by
virtual inheritance, dealing with the diamond problem by keeping only
one copy of the base class~\cite{ellis1990annotated}. However it 
suffers from object initialization problem~\cite{malayeri2009cz}.
It bypasses constructor calls to virtual superclasses, which can
cause serious semantic errors. In our approach, the \mixin
annotation has full control over object initialization,
and the mechanism is transparent to users. Moreover, if users are not satisfied with the default
 \texttt{of} method, they can provide customized factory methods.
